%!TEX root = ../thesis.tex
\begin{acknowledgements}

Workong on the final version of this thesis is a fantastic feeling, but it also marks the official end of a great period in my life. I spent four perfect years as a member of the Computational and Biological Learning Lab in Cambridge. I want to use this space to thank everybody who contributed to my feeling this way.

CBL is the best place to do a PhD in machine learning, academically and socially. It was M\'{a}t\'{e} Lengyel who got me started on Bayesian statistics and machine learning and having a mentor and friend like him around has been invaluable ever since. Zoubin and Carl are best in their field. Their excellent summer school lectures inspired me to apply for a PhD in Cambridge, and I have learnt so much from them since then.

I am lucky to have joined CBL early enough to know the very first batch of PhD students. This also meant I was the first generation to face the experience of people graduating and leaving. I wish the stimulating environment that CBL have built will last and inspire several future generations of students.

Completing my thesis research would not have been possible without my close collaborators and coauthors Neil, David, Miguel, Simon, Mohammad, Gerg\H{o}, Daniel, M\'{a}t\'{e} and Zoubin. Thanks to Phil Dawid for useful discussions about scoring rules. Thanks to all the anonymous reviewers of numerous conference reviews and rejection letters over the years; no offence taken. Thanks to Mischa, Sid, Jurgen, Neil, David and Miguel for help with finishing my thesis. 

I am grateful for Trinity College for funding my studies, and I am proud to have been a member of this great institution. Finally, I want to thank PeerIndex for their support and patience in the final struggle to write up my thesis.

Man\'{o}, D\'{a}vid, Lil\'{o}, Samu, Zitti. Cambridge would not have been the same without my friends and their families. Or indeed without my family, and Emese in particular. Thanks to you all.

\end{acknowledgements}